Linking assumptions and number of response options modulate inferred scalar implicature rate

### 2018-05-09 writeup.Rmd
addressed comments

The most prominently studied phenomenon in experimental pragmatics is undoubtedly *scalar implicature*. Scalar implicatures arise in virtue of **bw: as a result of?** a speaker producing the weaker of two ordered scalemates [@grice1975; @Horn1972; @Hirschberg1985; @Geurts2010]. Examples are provided in (1-2).

@@

A standard measure that has stood proxy **bw: stood as a proxy** for implicature rate across many studies is the proportion of 'pragmatic' judgments in truth-value judgment paradigms [@Bott2004;@Noveck2001;@Noveck2003;@Chemla2011;@Geurts2009;@DegenTanenhaus2015;@DeNeys2007;@Degen2014]. In these kinds of tasks, participants are provided a set of facts, either presented visually or via their own knowledge of the world. They are then asked to judge whether a sentence intended to describe those facts is true or false (or alternatively, whether it is right or wrong, or they are asked whether they agree or disagree with the sentence). The crucial condition for assessing implicature rates in these kinds of studies typically consists of a case where the facts are such that the stronger alternative is true and the target utterance is thus also true but underinformative. For instance, @Bott2004 asked participants to judge sentences like 'Some elephants are mammals', when world knowledge dictates that all elephants are mammals. Similarly, @DegenTanenhaus2015 asked participants to judge sentences like 'You got some of the gumballs' in situations where the visual evidence indicated that the participant received all the gumballs from a gumball machine. In these kinds of scenarios, the story goes, if a participant responds 'FALSE', that indicates that they computed a scalar implicature, eg to the effect of 'Not all elephants are mammals' or 'You didn't get all of the gumballs', which is (globally or contextually) false. If instead a participant responds 'TRUE', that is taken to indicate that they interpreted the utterance literally as `Some, and possibly all, elephants are mammals' or 'You got some, and possibly all, of the gumballs'.

@@

We begin by reporting an experiment that addresses the following question: do the number of response options provided in a truth-value judgment task and the way that responses are grouped into pragmatic ('SI') and literal ('no SI') change inferences about scalar implicature rates? Note that this way of asking the question presupposes two things: first, that whatever participants are doing in a truth-value judgment task, the behavioral measure can be interpreted as providing a measure of *interpretation*. And second, **bw: change to semicolon, i.e. "... interpretation; and second,**  that listeners either do or do not compute an implicature on any given occasion. In the General Discussion we will discuss both of these issues. First, following @Degen2014, we will offer some remarks on why truth-value judgment tasks are better thought of as measuring participants' estimates of speakers' *production* probabilities. This will suggest a completely different class of linking functions. And second, **bw: Next, (to avoid redundancy -- see above)** we discuss an alternative conception of scalar implicature as a probabilistic phenomeonen, a view that has recently rose to prominence in the subfield of probabilistic pragmatics [@Franke2016;@Goodman2016]. This alternative conception of scalar implicature, we argue, affords developing and testing quantitative linking functions in a rigorous and motivated way.

@@

Let's examine an example. Assume three response options (wrong, neither, right). Assume further that a third of participants each gave each of the three responses, **bw: Assume further that each of the three responses was selected by a third of participants,** i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are for instance **bw: if we are, for instance, **interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are. In the experiment reported in the following section, we presented participants with exactly this setup. We manipulated the number of response options between participants and analyzed the results under different linking assumptions.

@@

200 participants were recruited via Amazon Mechanical Turk. They optionally provided demographic information at the end of the study. Participants' mean age was `r round(mean_age)`. We also asked participants if they had any prior training in logic. `r logicTraining[[2]]` participants reported that they did, while `r logicTraining[[1]]` had no prior training in logic. **JD: Do we do any analysis of logic training???** All participants' data was included in the final analysis.

@@

The study was administered online through Amazon Mechanical Turk.[^2] Participants were first introduced to the set of cards we used in the study (Figure \@ref(fig:stimuli)). Each card depicted one or two animals, where an animal could be either a cat, a dog, or an elephant. Then participants were introduced to a blindfolded fictional character called Bob. Bob was blindfolded to avoid violations of ignorance expectations associated with the use of disjunction [@???] **JD: Masoud, can you insert citation pls? Geurts, Sauerland?**. Participants were told that Bob would guess the contents of the cards and their task was to indicate whether Bob's guess was wrong or right. On each trial, participants saw a card and a sentence representing Bob's guess. For example, they saw a card with a cat and read the sentence "There is a cat on the card." They then provided an assessment of Bob's guess.  The study ended after 24 trials.

[^2]: The experiment can be viewed [here](https://github.com/thegricean/si-paradigms/). **JD: Masoud, please turn this into a working link to the actual experiment**

@@

To analyze the effect of link and response options on inferred implicature rate, we used a Bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms] with uninformative priors **JD: Masoud is the uninformative prior part correct?**. The model predicted the log odds of implicature over no implicature from fixed effects of *response type* (binary, ternary, quaternary, quinary -- dummy-coded with binary as reference level), *link* (strong vs. weak -- dummy-coded with strong as reference level), and trial type (exhaustive vs. scalar -- dummy-coded, with exhaustive as reference level), as well as their two-way and three-way interactions. Following @barr2013random, we included the maximal random effects structure justified by the design: random intercepts for items (cards) and participants, random by-participant slopes for link, trial type, and their interaction, and random by-item slopes for link, trial type, response type, and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^3].

The model provided evidence for the following effects: First, there was a main effect of trial type such that scalar trials resulted in greater implicature rates than exhaustive trials (Mean Estimate = `r round(fixed_effects$Estimate[6],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[6],2)`, `r round(fixed_effects$X97.5.ile[6],2)`]). Second, there was an interaction between link and number of response options such that the quaternary task (Mean Estimate = `r round(fixed_effects$Estimate[7],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[7],2)`, `r round(fixed_effects$X97.5.ile[7],2)`]) and the quinary task  (Mean Estimate = `r round(fixed_effects$Estimate[8],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[8],2)`, `r round(fixed_effects$X97.5.ile[8],2)`]) with a weak link resulted in greater implicature rates. Finally, there was a three-way interaction between link, trial type, and number of response options (Mean Estimate = `r round(fixed_effects$Estimate[15],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[15],2)`, `r round(fixed_effects$X97.5.ile[15],2)`]). One interpretation of this interaction is that the difference between the weak and strong link on scalar trials in the quinary task was smaller than on exhaustive trials, though we believe this is not too interesting, given that the binary reference level implicature estimate was lower for exhaustive trials in the first place. **JD: happy to throw out this last sentence** Crucially, both number of response options and link affect the inferred implicature rate.

@@

## Summary and methodological discussion

@@

## Revisiting linking assumptions

On the traditional view of the link between implicature and behavior in sentence verification tasks, scalar implicature is conceptualized as a binary, categorical affair - that is, an implicature is either ‘calculated’ or it isn’t, and the behavioral reflexes of this categorical interpretation process should be straightforwardly observed in experimental paradigms. This assumption raises concerns for analyzing variation in behavior on a truth value judgment task; for example, why did the majority of respondents in the binary condition of our experiment answer “right” to an utterance of the underinformative "There is a cat or dog" when the card had both a cat and a dog on it? And why did a sizeable minority nonetheless choose "wrong" in this same condition?

To explain these data on the traditional view, we are forced to say that a) not all participants calculated the implicature; or that b) some participants who calculated the implicature did not choose the anticipated (i.e., "wrong") response due to some other cognitive process which overrode the 'correct' implicature behavior; or some mixture of (a) and (b). We might similarly posit that one or both of these factors underlie the variation in the ternary, quaternary, and quinary conditions. However, without an understanding of how to quantitatively specify the link between implicature calculation and its behavioral expression, the best we can hope for on this approach is an analysis which predicts general qualitative patterns in the data (e.g. a prediction of relatively more "right" responses than "wrong" responses in a given trial of our binary truth-value judgment task, or a prediction of a rise in the rate of response of "right"/"wrong" between two experimental conditions, given some contextual manipulation). However, we should stress that to the best of our knowledge, even a qualitative analysis of this kind of variation in behavior on sentence verification tasks - much less the effect of the number of response choices on that behavior - is largely underdeveloped scalar implicature literature.

We contrast the above view of implicature and its behavioral reflexes with an alternative linking hypothesis. Recent developments in the field of probabilistic pragmatics have demonstrated that pragmatic production and comprehension can be captured within the Rational Speech Act (RSA) framework  [@Frank2012; @Goodman2013; @Goodman2016; @Franke2016; @Qing2015; @Degen2013; @DegenTG2015; @Bergen2016; @Kao2014; @Scontras2017]. Much in the spirit of Gricean approaches to pragmatic competence, the RSA framework takes as its point of departure the idea that individuals are rational, goal-oriented communicative agents, who in turn assume that their interlocutors similarly behave according to general principles of cooperativity in communication. Just as in more traditional Gricean pragmatics, pragmatic inference and pragmatically-cooperative language production in the RSA framework are, at their core, the product of counterfactual reasoning about alternative utterances that one might produce (but does not, in the interest of cooperativity). However, the RSA framework explicitly and quantitatively models cooperative interlocutors as agents whose language production and comprehension is a function of Bayesian probabilistic inference regarding other interlocutors' expected behavior in a discourse context.

@@

This view contrasts with the traditional view in that it is rooted in a quantitative formalization of pragmatic competence which provides us a continuous measure of pragmatic reasoning. In the RSA framework, individuals never categorically draw (or fail to draw) pragmatic inferences about the utterances they hear. For example, exclusivity readings of disjunction are represented in RSA as relatively lower posterior conditional probability of a conjunctive meaning on the $P_L$ distribution given an utterance of "or", compared to the prior probability of that meaning. Thus, absent auxiliary assumptions about what exactly would constitute 'implicature', it is not even possible to talk about rate of implicature calculation in the RSA framework. The upshot, as we show below, is that this view of pragmatic competence does allow us to talk explicitly and quantitatively about rates of observed behavior in sentence verification tasks.

@@

A main desideratum of a behavioral linking hypothesis given the RSA view of pragmatic competence is to transform continuous probability distributions into categorical outputs (e.g. responses of “right”/”wrong” in the case of the binary condition of our experiment). For a given utterance $u$ and an intended communicated meaning $s$, $S_1$(u | s) outputs a conditional probability of $u$ given $s$. For example, in the binary condition of our experiment where a participant evaluated "There is a cat or a dog" when there were both animals on the card, the participant has access to the mental representation of $S_1$ and hence to the $S_1$ conditional probability of producing the utterance "cat or dog"" given a dog and cat card: $S_1$("cat or dog" | cat and dog). According to the linking hypothesis advanced here, the participant provides a particular response to $u$ if the RSA speaker probability of $u$ lies within a particular probability interval, given an observed state of the world (i.e. the configuration of animals on the card in our experiment). We model a responder, $R$, who in the binary condition responds “right” to an utterance $u$ in world $s$ just in case $S_1(u | s)$ exceeds some probability threshold $\theta$:

@@

= “right” iff $S_1$(u | s) $>$ $\theta$

= “wrong” otherwise

Because $\theta$ is a free parameter in our model, we have a straightforward understanding of between-subject response variation on a given trial of our experiment according to this linking hypothesis. **JD: Hold on, Brandon, what's the link between being a free parameter hand having an understanding of between-subject response variation?** That is, we can attribute this variation to individual differences vis a vis the value of the probability threshold against which utterances are evaluated in a given world state. Moreover, the model of a responder in the binary condition is extended intuitively to the condition where participants had three response options. In this case, we allow for two probability thresholds: $\theta_1$, the minimum standard for an utterance in a given world state to count as "right", and $\theta_2$, the minimum standard for "neither". Thus, in the ternary condition, R(u, s, $\theta_1$ , $\theta_2$) is “right” iff $S_1$(u | s) > $\theta_1$ and “neither” iff $\theta_1$ > $S_1$(u | s) > $\theta_2$.
To fully generalize the model to our five experimental conditions, we say that $R$ takes as its input an utterance $u$, a world state $s$, and a number of threshold variables dependent on a variable $c$, corresponding to the experimental condition in which the participant finds themself (e.g. the range of possible responses available to $R$).

@@

= “right” iff $S_1$(u | s) $>$ $\theta_1$

= “neither” iff $\theta_1$ $>$ $S_1$(u | s) $>$ $\theta_2$

= “wrong” otherwise

Given c = “quaternary”

@@

= “right” iff $S_1$(u | s) $>$ $\theta_1$

= “kinda right” iff $\theta_1$ $>$ $S_1$(u | s) $>$ $\theta_2$

= “kinda wrong” iff $\theta_2$ $>$ $S_1$(u | s) $>$ $\theta_3$

= “wrong” otherwise

@@

= “right” iff $S_1$(u | s) $>$ $\theta_1$

=“kinda right” iff $\theta_1$ $>$ $S_1$(u | s) $>$ $\theta_2$

= “neither” iff $\theta_2$ $>$ $S_1$(u | s) $>$ $\theta_3$

= “kinda wrong” iff $\theta_3$ $>$ $S_1$(u | s) $>$ $\theta_4$

= “wrong” otherwise

The above analysis is a proof of concept for the following idea: by relaxing the assumptions of the traditional view of scalar implicature (namely, that scalar implicatures either are or are not calculated, and that behavior on sentence verification tasks directly reflects this binary interpretation process), we can propose quantitative models of the variation in behavior we observe in experimental settings. **JD: Brandon, I feel what's still missing here is a demonstration that this is a _reasonable_ linking function. Maybe we can add the point that with this linking function, there's nothing inherently special about the implicature conditions -- the same linking function can work for any of the eight conditions in our experiment, because they're all the result of a participant having to make a behavioral choice in the face of an underlying speaker probability.**

However, this way of understanding behavior in truth value judgment tasks also raises new analytic and empirical questions which we must leave to future work. For instance, given the experimental results reported here, what is the range of plausible values for, e.g., the single probability threshold $\theta$, which separates "right" from "wrong" utterances in the binary condition? Bayesian statistical methods provide us the means for estimating the value of $\theta$ in our RSA-based model, by allowing us to recover a posterior distribution of possible values of $\theta$ given an assumption of a uniform prior and our empirically-observed data, i.e., the actual rate of response of "right" on each trial of the binary condition of the experiment. We may also ask, for example, whether the speaker probability threshold for an utterance being "right" is relatively constant across conditions or changes systematically with the introduction of more response options. Again, Bayesian data analysis allows us to explore this question by investigating the extent to which the inferred posterior distribution of, e.g., possible values of the threshold for "right" in the binary condition of the experiment differs from the inferred posterior distribution of possible values of the threshold for "right" in the ternary, quatenary, and quinary conditions.

@@

**JD: We need a final concluding paragraph that reiterates the main points.**

>>>>

The most prominently studied phenomenon in experimental pragmatics is undoubtedly *scalar implicature*. Scalar implicatures arise in virtue of a speaker producing the weaker of two ordered scalemates [@grice1975; @Horn1972; @Hirschberg1985; @Geurts2010]. Examples are provided in (1-2).

@@

A standard measure that has stood proxy for implicature rate across many studies is the proportion of 'pragmatic' judgments in truth-value judgment paradigms [@Bott2004;@Noveck2001;@Noveck2003;@Chemla2011;@Geurts2009;@DegenTanenhaus2015;@DeNeys2007;@Degen2014]. In these kinds of tasks, participants are provided a set of facts, either presented visually or via their own knowledge of the world. They are then asked to judge whether a sentence intended to describe those facts is true or false (or alternatively, whether it is right or wrong, or they are asked whether they agree or disagree with the sentence). The crucial condition for assessing implicature rates in these kinds of studies typically consists of a case where the facts are such that the stronger alternative is true and the target utterance is thus also true but underinformative. For instance, @Bott2004 asked participants to judge sentences like 'Some elephants are mammals', when world knowledge dictates that all elephants are mammals. Similarly, @DegenTanenhaus2015 asked participants to judge sentences like 'You got some of the gumballs' in situations where the visual evidence indicated that the participant received all the gumballs from a gumball machine. In these kinds of scenarios, the story goes, if a participant responds 'FALSE', that indicates that they computed a scalar implicature, eg to the effect of 'Not all elephants are mammals' or 'You didn't get all of the gumballs', which is (globally or contextually) false. If instead a participant responds 'TRUE', that is taken to indicate that they interpreted the utterance literally as `Some, and possibly all, elephants are mammals' or 'You got some, and possibly all, of the gumballs'.

@@

We begin by reporting an experiment that addresses the following question: do the number of response options provided in a truth-value judgment task and the way that responses are grouped into pragmatic ('SI') and literal ('no SI') change inferences about scalar implicature rates? Note that this way of asking the question presupposes two things: first, that whatever participants are doing in a truth-value judgment task, the behavioral measure can be interpreted as providing a measure of *interpretation*. And second, that listeners either do or do not compute an implicature on any given occasion. In the General Discussion we will discuss both of these issues. First, following @Degen2014, we will offer some remarks on why truth-value judgment tasks are better thought of as measuring participants' estimates of speakers' *production* probabilities. This will suggest a completely different class of linking functions. And second, we discuss an alternative conception of scalar implicature as a probabilistic phenomeonen, a view that has recently rose to prominence in the subfield of probabilistic pragmatics [@Franke2016;@Goodman2016]. This alternative conception of scalar implicature, we argue, affords developing and testing quantitative linking functions in a rigorous and motivated way.

@@

Let's examine an example. Assume three response options (wrong, neither, right). Assume further that a third of participants each gave each of the three responses, i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are for instance interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are. In the experiment reported in the following section, we presented participants with exactly this setup. We manipulated the number of response options between participants and analyzed the results under different linking assumptions.

@@

200 participants were recruited via Amazon Mechanical Turk. They optionally provided demographic information at the end of the study. Participants' mean age was `r round(mean_age)`. We also asked participants if they had any prior training in logic. `r logicTraining[[2]]` participants reported that they did, while `r logicTraining[[1]]` had no prior training in logic. All participants' data was included in the final analysis.

@@

The study was administered online through Amazon Mechanical Turk.[^2] Participants were first introduced to the set of cards we used in the study (Figure \@ref(fig:stimuli)). Each card depicted one or two animals, where an animal could be either a cat, a dog, or an elephant. Then participants were introduced to a blindfolded fictional character called Bob. Bob was blindfolded to avoid violations of ignorance expectations associated with the use of disjunction [@chierchia2001;@sauerland2004scalar] Participants were told that Bob would guess the contents of the cards and their task was to indicate whether Bob's guess was wrong or right. On each trial, participants saw a card and a sentence representing Bob's guess. For example, they saw a card with a cat and read the sentence "There is a cat on the card." They then provided an assessment of Bob's guess.  The study ended after 24 trials.

[^2]: The experiment can be viewed [here](https://cdn.rawgit.com/thegricean/si-paradigms/94a590f0/experiments/main/1_methods/online_experiment/connective_game.html).

@@

To analyze the effect of link and response options on inferred implicature rate, we used a Bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms] with uninformative or weakly informative priors [^3]. The model predicted the log odds of implicature over no implicature from fixed effects of *response type* (binary, ternary, quaternary, quinary -- dummy-coded with binary as reference level), *link* (strong vs. weak -- dummy-coded with strong as reference level), and trial type (exhaustive vs. scalar -- dummy-coded, with exhaustive as reference level), as well as their two-way and three-way interactions. Following @barr2013random, we included the maximal random effects structure justified by the design: random intercepts for items (cards) and participants, random by-participant slopes for link, trial type, and their interaction, and random by-item slopes for link, trial type, response type, and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^3].

[^3: To know more about the default priors of the "brms" package you can refer to [the brms package manual](ftp://cran.r-project.org/pub/R/web/packages/brms/brms.pdf).]

The model provided evidence for the following effects: First, there was a main effect of trial type such that scalar trials resulted in greater implicature rates than exhaustive trials (Mean Estimate = `r round(fixed_effects$Estimate[6],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[6],2)`, `r round(fixed_effects$X97.5.ile[6],2)`]). Second, there was an interaction between link and number of response options such that the quaternary task (Mean Estimate = `r round(fixed_effects$Estimate[7],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[7],2)`, `r round(fixed_effects$X97.5.ile[7],2)`]) and the quinary task  (Mean Estimate = `r round(fixed_effects$Estimate[8],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[8],2)`, `r round(fixed_effects$X97.5.ile[8],2)`]) with a weak link resulted in greater implicature rates. Finally, there was a three-way interaction between link, trial type, and number of response options (Mean Estimate = `r round(fixed_effects$Estimate[15],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[15],2)`, `r round(fixed_effects$X97.5.ile[15],2)`]). One interpretation of this interaction is that the difference between the weak and strong link on scalar trials in the quinary task was smaller than on exhaustive trials, though we believe this is not too interesting, given that the binary reference level implicature estimate was lower for exhaustive trials in the first place. Crucially, both number of response options and link affect the inferred implicature rate.

@@

## Discussion

@@

On the traditional view of the link between implicature and behavior in sentence verification tasks, scalar implicature is conceptualized as a binary, categorical affair - that is, an implicature is either ‘calculated’ or it isn’t, and the behavioral reflexes of this categorical interpretation process should be straightforwardly observed in experimental paradigms. This assumption raises concerns for analyzing variation in behavior on a truth value judgment task; for example, why did the majority of respondents in the binary condition of our experiment answer “Right” to an utterance of the underinformative "There is a cat or dog" when the card had both a cat and a dog on it? And why did a sizeable minority nonetheless choose "Wrong" in this same condition?

To explain these data on the traditional view, we are forced to say that a) not all participants calculated the implicature; or that b) some participants who calculated the implicature did not choose the anticipated (i.e., "Wrong") response due to some other cognitive process which overrode the 'correct' implicature behavior; or some mixture of (a) and (b). We might similarly posit that one or both of these factors underlie the variation in the ternary, quaternary, and quinary conditions. However, without an understanding of how to quantitatively specify the link between implicature calculation and its behavioral expression, the best we can hope for on this approach is an analysis which predicts general qualitative patterns in the data (e.g. a prediction of relatively more "Right" responses than "Wrong" responses in a given trial of our binary truth-value judgment task, or a prediction of a rise in the rate of response of "Right"/"Wrong" between two experimental conditions, given some contextual manipulation). **XXX Judith, I was thinking of adding the sentence to end this paragraph: "However, we should stress that to the best of our knowledge, even a qualitative analysis of this kind of variation in behavior on sentence verification tasks - much less the effect of the number of response choices on that behavior - is largely underdeveloped scalar implicature literature." But it might be way too strong a claim? **

**XXX Brandon, can you be more specific about what you mean by 'tracing a general qualitative pattern'? XXX Judith: I've tried to offer some examples of what I mean by this, see above.**

We contrast the above view of implicature and its behavioral reflexes with an alternative linking hypothesis. Recent developments in the field of probabilistic pragmatics have demonstrated that pragmatic production and comprehension can be captured within the Rational Speech Act (RSA) framework, which treats listeners and speakers as agents that reason about each other and choose interpretations and utterances, respectively, that soft-maximize utility [@Frank2012; @Goodman2013; @Goodman2016; @Franke2016; @Qing2015; @Degen2013; @DegenTG2015; @Bergen2016; @Kao2014; @Scontras2017]. **XXX Judith: re: your Slack comments, in the following few sentences I've provided a summary of the intuition behind RSA (going off of the discussion in Goodman & Stuhlmueller 2013)** Much in the spirit of Gricean approaches to pragmatic competence, the RSA framework takes as its point of departure the idea that individuals are rational, goal-oriented communicative agents, who in turn assume that their interlocutors similarly behave according to general principles of cooperativity in communication. Just as in more traditional Gricean pragmatics, pragmatic inference and pragmatically-cooperative language production in the RSA framework are, at their core, the product of counterfactual reasoning about alternative utterances that one might produce (but does not, in the interest of cooperativity). However, the RSA framework explicitly and quantitatively models coooperative interlocutors as agents whose language production and comprehension is a function of Bayesian probalistic inference regarding other interlocutors' expected behavior in a discourse context.

@@

Thus, our alternative linking hypothesis contrasts with the traditional view in that it is rooted in a quantitative formalization of pragmatic competence which provides us a continuous measure of pragmatic reasoning. In the RSA framework, individuals never categorically draw (or fail to draw) pragmatic inferences about the utterances they hear. For example, exclusivity readings of disjunction are represented in RSA as relatively lower posterior conditional probability of a conjunctive meaning on the $P_L$ distribution given an utterance of "or", compared the prior probability of that meaning. Thus, absent auxiliary assumptions about what exactly would constitute 'implicature', it is not even possible to talk about rate of implicature calculation in the RSA framework. The upshot, as we show below, is that this view of pragmatic competence does allow us to talk explicitly and quantitatively about rates of observed behavior in sentence verification tasks.

@@

A main desideratum of a behavioral linking hypothesis given the RSA view of pragmatic competence is to transform continuous probability distributions into categorical outputs (e.g. responses of “Right”/”Wrong” in the case of the binary condition of our experiment). For a given utterance $u$ and an intended communicated meaning $s$, $S_1$(u | s) outputs a conditional probability of $u$ given $s$. For example, in the binary condition of our experiment where a participant evaluated "There is a cat or a dog" when there were both animals on the card, the participant has access to the mental representation of $S_1$ and hence to the $S_1$ conditional probability of producing the utterance "cat or dog"" given a dog and cat card: $S_1$("cat or dog" | cat and dog). According to the linking hypothesis advanced here, the participant provides a particular response to $u$ if the RSA speaker probability of $u$ lies within a particular probability interval, given an observed state of the world (i.e. the configuration of animals on the card in our experiment). We model a responder, $R$, who in the binary condition responds “Right” to an utterance $u$ in world $s$ just in case $S_1(u | s)$ exceeds some probability threshold $\theta$:

@@

= “Right” iff $S_1$(u | s) $>$ $\theta$

= “Wrong” otherwise

**XXX The following paragraph is mostly new:** Because $\theta$ is a free parameter in our model, we have a straightforward understanding of between-subject response variation on a given trial of our experiment according to this linking hypothesis. That is, we can attribute this variation to individual differences vis a vis the value of the probability threshold against which utterances are evaluated in a given world state. Moreover, the model of a responder in the binary condition is extended intuitively to the condition where participants had three response options. In this case, we allow for two probability thresholds: $\theta_1$, the minimum standard for an utterance in a given world state to count as "Right", and $\theta_2$, the minimum standard for "Neither". Thus, in the ternary condition, R(u, s, $\theta_1$ , $\theta_2$) is “Right” iff $S_1$(u | s) > $\theta_1$ and “Neither” iff $\theta_1$ > $S_1$(u | s) > $\theta_2$.
To fully generalize the model to our five experimental conditions, we say that $R$ takes as its input an utterance $u$, a world state $s$, and a number of threshold variables dependent on a variable $c$, corresponding to the experimental condition in which the participant finds herself (e.g. the range of possible responses available to $R$).

@@

= “Right” iff $S_1$(u | s) $>$ $\theta_1$

= “Neither” iff $\theta_1$ $>$ $S_1$(u | s) $>$ $\theta_2$

= “Wrong” otherwise

Given c = “quatenary”

@@

= “Right” iff $S_1$(u | s) $>$ $\theta_1$

= “Kinda Right” iff $\theta_1$ $>$ $S_1$(u | s) $>$ $\theta_2$

= “Kinda Wrong” iff $\theta_2$ $>$ $S_1$(u | s) $>$ $\theta_3$

= “Wrong” otherwise

@@

= “Right” iff $S_1$(u | s) $>$ $\theta_1$

=“Kinda Right” iff $\theta_1$ $>$ $S_1$(u | s) $>$ $\theta_2$

= “Neither” iff $\theta_2$ $>$ $S_1$(u | s) $>$ $\theta_3$

= “Kinda Wrong” iff $\theta_3$ $>$ $S_1$(u | s) $>$ $\theta_4$

= “Wrong” otherwise

The above analysis is a proof of concept for the following idea: by relaxing the assumptions of the traditional view of scalar implicature (namely, that scalar implicatures either are or are not calculated, and that behavior on sentence verification tasks directly reflects this binary interpretation process), we can propose quantitative models of the variation in behavior we observe in experimental settings.

**XXX The following is entirely new:** However, this way of understanding behaivor in truth value jugdment tasks also raises new analytic and empirical questions which we must leave to future work. For instance, given the experimental results reported here, what is the range of plausible values for, e.g., the single probability threshold $\theta$, which separates "Right" from "Wrong" utterances in the binary condition? Bayesian statistical methods provide us the means for estimating the value of $\theta$ in our RSA-based model, by allowing us to recover a posterior distribution of possible values of $\theta$ given an assumption of a uniform prior and our empirically-observed data, i.e. the actual rate of response of "Right" On each trial of the binary condition of the experiment. We may also ask, for example, whether the speaker probability threshold for an utterance being "Right" is relatively constant across conditions or changes systematically with the introduction of more response options. Again, Bayesian data analysis allows us to explore this question by investigating the extent to which the inferred posterior distribution of, e.g., possible values of the threshold for "Right" in the binary condition of the experiment differs from the inferred posterior distribution of possible values of the threshold for "Right" in the ternary, quatenary, and quinary conditions.

@@
